---
title: "course project 1"
author: "Yan He, FRM"
date: "10/13/2019"
output:
    prettydoc::html_pretty:
        theme: cayman
        highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data

At first we load the activity data.

```{r load data}
library(tidyverse)

setwd('~/coursera_projects/reproducible_research')
activity = read.csv('activity.csv', header = TRUE)
summary(activity)
```

There are three variables in this dataset: 

* steps: Number of steps taking in a 5-minute interval;
* date: The date on which the measurement was taken in YYYY-MM-DD format
* interval: Identifier for the 5-minute interval in which measurement was taken

## Processing and Plotting Data

### 1. Histogram of the total number of steps taken each day

```{r plots, echo=TRUE,warning=FALSE}
## Data Transform

activity$date = as.Date(activity$date)


sum_steps_by_dt = activity %>% group_by(date) %>% summarise(sum_steps = sum(steps))

steps_hist = ggplot(data = sum_steps_by_dt, aes(x = sum_steps, y = ..count..))+geom_histogram()
steps_hist

png(filename = './figure/plot1.png')
ggplot(data = sum_steps_by_dt, aes(x = sum_steps, y = ..count..))+geom_histogram()
dev.off()
```

### 2. Mean and median number of steps taken each day

```{r statistics, echo = TRUE}
mean_of_steps = mean(sum_steps_by_dt$sum_steps, na.rm = TRUE)
mean_of_steps

median_of_steps = median(sum_steps_by_dt$sum_steps, na.rm = TRUE)
median_of_steps
```

### 3. Time series plot of the average number of steps taken
```{r time series, echo = TRUE}
mean_steps_by_dt = activity %>% group_by(date) %>% summarise(mean_of_steps = mean(steps))

ggplot(data = mean_steps_by_dt, aes(x = date, y = mean_of_steps)) + geom_line()

png(filename = './figure/plot2.png')
ggplot(data = mean_steps_by_dt, aes(x = date, y = mean_of_steps)) + geom_line()
dev.off()
```

### 4. The 5-minute interval that, on average, contains the maximum number of steps
```{r, echo = TRUE}
mean_steps_by_interval = activity %>% group_by(interval) %>% summarise(mean_steps = mean(steps, na.rm = TRUE)) 

max(mean_steps_by_interval$mean_steps)

```

### 5. Code to describe and show a strategy for imputing missing data

This easiest way to imput missing data is to remove them.

```{r na data, echo= TRUE}
if(sum(is.na(activity) == TRUE)>0){
    no_missing_activity = activity[complete.cases(activity),]
}
```

### 6. Histogram of the total number of steps taken each day after missing values are imputed
```{r hist, echo = TRUE}

sum_steps_by_dt_new = no_missing_activity %>% group_by(date) %>% summarise(sum_steps = sum(steps))

steps_hist_new = ggplot(data = sum_steps_by_dt_new, aes(x = sum_steps, y = ..count..))+geom_histogram()
steps_hist_new

png(filename = './figure/plot3.png')
steps_hist_new
dev.off()
```

### 7. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r panel plot, echo = TRUE}
activity_by_weekdays = no_missing_activity %>% 
     mutate(day_of_week = weekdays(date)) %>%
     mutate(weekend = ifelse(day_of_week  == 'Saturday' | day_of_week == 'Sunday','Weekend','Weekday'))

activity_by_weekdays %>% 
     group_by(weekend, interval) %>% 
     mutate(mean_steps_by_interval = mean(steps)) %>% 
     ggplot(aes(x = interval, y = mean_steps_by_interval)) + geom_line() + facet_wrap(~weekend)

png(filename = './figure/plot4.png')
activity_by_weekdays %>% 
     group_by(weekend, interval) %>% 
     mutate(mean_steps_by_interval = mean(steps)) %>% 
     ggplot(aes(x = interval, y = mean_steps_by_interval)) + geom_line() + facet_wrap(~weekend)
dev.off()
```